---
title: "Stats 506 Midterm Project: Hosmer-Lemeshow Model Calibration"
author: "Group 5: Hao He, Yanhan Si"
date: "`r format.Date(Sys.Date(), '%B %d, %Y')`"
output: html_document
     
---

```{r setup, include=FALSE}
# 79: -------------------------------------------------------------------------
knitr::opts_chunk$set(echo = FALSE)
```

# Introduction

The purpose of this tutorial is to introduce the Hosmer-Lemeshow Test, which is
a goodness of fit test for logistic regression models. We will first introduce
the key concepts behind logistic regression models and the Hosmer-Lemeshow Test.
Then we will use a bank churn data to illustrate how to perform Hosmer-Lemeshow
test and create Hosmer-Lemeshow plot in **R** and **Stata**. Eventually we will
compare the result to a likelihood based approach like AIC and explain the
potential limitations with the HL test.

### Logistic Regression Model

Logistic regression is a statistical model that in its basic form uses a
logistic function to model a binary response variable. In our example, the
response variable will be whether the customer churn or not. Let's denote the
response variable to have value 0 or 1 in our case.
The logistic model predicts $P(Y=1) = \frac{1}{1+exp(\beta'x)}$, so $P(Y=0) = \frac{exp(\beta'x)}{1+exp(\beta'x)}$.

### Hosmer-Lemeshow Test

The Hosmer-Lemeshow Test assesses whether or not the observed event rates match
expected event rates in subgroups of the model population. The Hosmer–Lemeshow
test statistic is given by:  
$$H = \sum_{g=1}^{G}\frac{(O_{1g}-E_{1g})^2}{N_g\pi_g(1-\pi_g)}$$  
, where $O_{1g}$, $E_{1g}$, $O_{0g}$, $E_{0g}$, $N_g$, and $π_g$ denote the
observed $Y=1$ events, expected $Y=1$ events, observed $Y=0$ events, expected
$Y=0$ events, total observations, predicted risk for the $g^{th}$ risk decile
group, and G is the number of groups. The test statistic asymptotically follows
a $\chi ^{2}$ distribution with G − 2 degrees of freedom. For the HL test, the
null hypothesis is that the observed and expected proportions are the same
across subgroups, which indicates a good fit model.

### Data Description
We will use the bank churn data available on Kaggle([ Data Source](https://www.kaggle.com/shrutimechlearn/churn-modelling)).
The data contains 10000 observations of 14 variables of which 2000 are postivie observations. For simplicity, we only
construct the logistic model using the variables below:  
`Credit Score`: The credit score of the customer, ranging from 350 to 850  
`Geography`: The country where the customer belongs. There are only three
countries in our case, which are Germany, France and Spain.  
`Gender`: Gender  
`Age`: Age   
`Tenure`: Number of years for which the customer has been with the bank  
`Balance`: Bank balance of the customer  
`HasCrCard`: Binary Flag for whether the customer holds a credit card with the
bank or not.  
`IsActiveMember`: Binary Flag for whether the customer is an active member with
the bank or not.  
`EstimatedSalary`: Estimated salary of the customer in Dollars.  
`Exited`: The response variable indicated whether the customer churn or not. 1
means customer exited and 0 means retained.

# Data Visualization

```{r}
library(tidyverse)
library(ggplot2)
library(readr)
library(ResourceSelection)
# data: -----------------------------------------------------------------------
path = "./data"
file_name = sprintf('%s/Churn_Modelling_Cleaned.csv', path)
data = read_csv(file_name)
# factoring: ------------------------------------------------------------------
names(data)
data = data %>% mutate(Exited = as.factor(Exited),
                       HasCrCard = as.factor(HasCrCard),
                       IsActiveMember = as.factor(IsActiveMember))
summary(data)
head(data)
data %>% ggplot(aes( CreditScore,color = Exited )) + geom_density()
```

```{r}
data %>% ggplot(aes(x = Exited, y = Age, color = Gender)) + geom_boxplot()
```

```{r}
data %>% ggplot(aes(x = Geography, y = Exited)) + geom_count()
```

```{r}
data %>% ggplot(aes(x = Exited, y = HasCrCard, color = Gender)) +geom_count()
```


# Model fitting
```{r}
data = data %>% mutate(Exited = as.factor(Exited),
                       HasCrCard = as.factor(HasCrCard),
                       IsActiveMember = as.factor(IsActiveMember))
# modeling: -------------------------------------------------------------------
model1 = glm(Exited ~ ., data = data, family = binomial(link="logit"))
summary(model1)
# Hosmer-Lemeshow Goodness of Fit Test: ---------------------------------------
hoslem.test(model1$y, model1$fitted.values, g = 10)
```



# References
[Logistic Model](https://en.wikipedia.org/wiki/Logistic_regression)  
[Hosmer-Lemeshow Test](https://en.wikipedia.org/wiki/Hosmer%E2%80%93Lemeshow_test#Pearson_chi-squared_goodness_of_fit_test)





